---
title: "knitted hw2"
author: "Roxanne Felig"
date: "2/8/2021"
output:
    html_document:
        theme: paper
    github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gapminder) # BMW: Put your library() calls in the setup chunk
library(tidyverse)
library(dplyr) # BMW: library(tidyverse) loads dplyr, so you don't need both.
```

## Assignment 02: Explore gapminder and use dplyr

### 1.1 Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.

<!-- BMW: In markdown, you need a blank line between paragraphs. A single line break is ignored. -->

```{r filter}
gapminder %>%
  filter(country %in% c("Argentina","Chile","Germany") & year>1969 & year<1980)
```

### 1.2 Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.

<!-- BMW: YOu need a space after the # for headings. -->

```{r select}
gapminder %>%
  filter(country %in% c("Argentina","Chile","Germany") & year>1969 & year<1980) %>% 
  select(country,gdpPercap)
```

### 1.3 Make a new variable in gapminder for the change in life expectancy from the previous measurement. Filter this table to show all of the entries that have experienced a drop in life expectancy.

Hint: you might find the lag() or diff() functions useful.

```{r new variable}
gapminder %>%
  group_by(country) %>% 
  mutate(lifeExp_change = lifeExp - lag(lifeExp)) %>% 
  filter(lifeExp_change < 0) # BMW: Great work! 
```

### 1.4 Filter gapminder so that it shows the max GDP per capita experienced by each country.

Hint: you might find the max() function useful helpful

```{r maxGDP}
gapminder %>%
  group_by(country) %>% 
  filter(maxgdp == max(gdpPercap)) # BMW: The question asked you to _filter_ (e.g., so that we can see which year had the max gdp)
```

### 1.5 Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot() function. In your plot, put GDP per capita on a log scale.

``` {r plot}
gapminder %>%
  filter(country == 'Canada') %>% # BMW: For readability, put spaces around = and other operators
  # select(gdpPercap,lifeExp) %>% # BMW: YOu don't need select(). Don't select unnecessarily
  ggplot(aes(x = gdpPercap, y = lifeExp)) + 
  geom_point() + 
  scale_x_log10() # BMW: For readability, put new plot layers on new lines

```

## Use palmerpenguins::penguins or another dataset of your choice. (Check out a dataset from the datasets R package if you want!)

### 2.1 Pick two quantitative variables to explore.

Make a summary table of descriptive statistics for these variables using summarize().
Include whatever statistics you feel appropriate (mean, median sd, range, etc.).
Make a scatterplot of these variables using ggplot().

__when i tried to pipe in the graph, it said it couldn't find the variables__

<!-- BMW: I'm not sure what you mean? You piped the data into the ggplot() function below. -->

```{r dataset exploring}
palmerpenguins::penguins %>% 
  summarize(
    mean_bill = mean(bill_length_mm, na.rm = TRUE),
    sd_bill = sd(bill_length_mm, na.rm = TRUE),
    mean_flip = mean(flipper_length_mm, na.rm = TRUE),
    sd_flip = sd(flipper_length_mm, na.rm = TRUE)
  )

palmerpenguins::penguins %>%
  ggplot(aes(x = bill_length_mm, y = flipper_length_mm)) +
  geom_point()
```

###2.2 Pick one categorical variable and one quantitative variable to explore.

Make a summary table giving the sample size (hint: n()) and descriptive statistics for the quantitative variable by group.
Make one or more useful plots to visualize these variables.

```{r categorical}
# str(palmerpenguins::penguins) # BMW: Don't include str() or View() or similar in your final script

palmerpenguins::penguins %>%
  group_by(island) %>%  # BMW: Don't select() unnecessarily. You also need to group by your categorical variable to get descriptives for each separately
  summarize(
    n = n(),
    mean_bill = mean(bill_length_mm, na.rm = TRUE),
    sd_bill = sd(bill_length_mm, na.rm=TRUE)
  ) 

palmerpenguins::penguins %>%
  ggplot(aes(x = island, y = bill_length_mm)) +
  # geom_bar(stat = 'identity') # geom_bar() isn't a good choice here. This plot is just showing you the total of the bill_length_mm for each island. That's not very useful. You probabaly want something like a scatter plot (geom_point or geom_jitter) and boxplot (geom_boxplot)
  geom_jitter(aes(color = island), height = 0, width = .2) +
  geom_boxplot(fill = "transparent")
```

```{r bonus}
filter(gapminder, country %in% c("Rwanda", "Afghanistan"))
##the first code skipped over several years by searching for anywhere where Rwanda was listed first, followed by Afghanistan.
```

<!-- BMW: Great job! -->
