---
title: "HW2"
author: "Roxanne Felig"
date: "2/7/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 02: Explore gapminder and use dplyr

```{r load packages}
# load your packages here:
library(gapminder)
library(tidyverse)
library(dplyr)
```

### 1.1 Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.
```{r filter}
gapminder %>%
  filter(country %in% c("Argentina","Chile","Germany") & year>1969 & year<1980)
```

###1.2 Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.

```{r select}
gapminder %>%
  filter(country %in% c("Argentina","Chile","Germany") & year>1969 & year<1980) %>% 
  select(country,gdpPercap)

```
###1.3 Make a new variable in gapminder for the change in life expectancy from the previous measurement. Filter this table to show all of the entries that have experienced a drop in life expectancy.
Hint: you might find the lag() or diff() functions useful.

```{r new variable}
gapminder %>%
  group_by(country) %>% 
  mutate(lifeExp_change=lifeExp-lag(lifeExp)) %>% 
  filter(lifeExp_change<0)

```

###1.4 Filter gapminder so that it shows the max GDP per capita experienced by each country.
Hint: you might find the max() function useful helpful

```{r maxGDP}
gapminder %>%
  group_by(country) %>% 
  summarize(maxgdp=max(gdpPercap))
  
```

### 1.5 Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot() function. In your plot, put GDP per capita on a log scale.

``` {r plot}
gapminder %>%
  select(gdpPercap,lifeExp) %>% 
  ggplot(aes(x=gdpPercap,y=lifeExp)) +geom_point() + scale_x_log10()

```

##Use palmerpenguins::penguins or another dataset of your choice. (Check out a dataset from the datasets R package if you want!)

###2.1 Pick two quantitative variables to explore.
Make a summary table of descriptive statistics for these variables using summarize().
Include whatever staistics you feel appropriate (mean, median sd, range, etc.).
Make a scatterplot of these variables using ggplot().

__when i tried to pipe in the graph, it said it couldn't find the variables__
```{r dataset exploring}
palmerpenguins::penguins %>% 
  select(bill_length_mm, flipper_length_mm) %>% 
  summarize(
    mean_bill=mean(bill_length_mm, na.rm=TRUE),
    sd_bill=sd(bill_length_mm, na.rm=TRUE),
    mean_flip=mean(flipper_length_mm, na.rm=TRUE),
    sd_flip=sd(flipper_length_mm, na.rm=TRUE))
palmerpenguins::penguins %>% 
  ggplot(aes(x=bill_length_mm,y=flipper_length_mm)) +geom_point()

```

###2.2 Pick one categorical variable and one quantitative variable to explore.
Make a summary table giving the sample size (hint: n()) and descriptive statistics for the quantitative variable by group.
Make one or more useful plots to visualize these variables.

```{r categorical}
str(palmerpenguins::penguins)
palmerpenguins::penguins %>% 
  select(island, bill_length_mm) %>% 
  summarize(n=n(),
            mean_bill=mean(bill_length_mm, na.rm=TRUE),
            sd_bill=sd(bill_length_mm, na.rm=TRUE)) 

palmerpenguins::penguins %>% 
  select(island, bill_length_mm) %>% 
  ggplot(aes(x=island,y=bill_length_mm)) +
  geom_bar(stat='identity')

```

```{r bonus}
filter(gapminder, country %in% c("Rwanda", "Afghanistan"))
##the first code skipped over several years by searching for anywhere where Rwanda was listed first, followed by Afghanistan.
```