---
title: "HW3"
author: "Roxanne Felig"
date: "2/8/2021"
output:
    html_document:
        theme: paper
    github_document: default
---
# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r}
dictionary <- readr::read_csv("/Users/roxannefelig/Desktop/progdata/hw-roxannefelig/HW3/green_dictionary.csv")
green_data <- readr::read_csv("/Users/roxannefelig/Desktop/progdata/hw-roxannefelig/HW3/green_data.csv")
```

```{r view files}
View(dictionary)
View(green_data)
library(tidyverse)
library(dplyr)
library(psych)
library(naniar)
library(reshape2)
```


```{r question 1}
str(green_data)
summary(green_data) #I notice that some missing values are coded as NA while others are coded as -99, so I'm going to re-code them

data<-green_data %>% 
  mutate_at(vars(comp1:student),
         ~ifelse(. == -99, NA, .)
  )

summary(data) #now we see that the min values for variables are 1, not -99.

plot(data$green1)

melt_data<-melt(data) #collapses data to one column
View(melt_data)
#plotting using the melted data and the facet wrap will allow us to see histograms for each scale item, and the frequency of each response choice.

ggplot(melt_data)+
  aes(x=value)+
  geom_histogram()+
  facet_wrap(~variable, 
             scales="free")
```


For your assignment, do the following.

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
     Is anything unusual?
  2. Compute total scores for the four scales. 
     Recode variables as needed.
  3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the recaled variables `*_pomp`.
  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?
  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
