---
title: "HW2"
author: "Roxanne Felig"
date: "2/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 02: Explore gapminder and use dplyr

```{r load packages}
# load your packages here:
library(gapminder)
library(tidyverse)
library(dplyr)
```

### 1.1 Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.
```{r filter}
gapminder %>%
  filter(country %in% c("Argentina","Chile","Germany") & year>1969 & year<1980)
```

###1.2 Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.

```{r select}
gapminder %>%
  filter(country %in% c("Argentina","Chile","Germany") & year>1969 & year<1980) %>% 
  select(country,gdpPercap)

```
###1.3 Make a new variable in gapminder for the change in life expectancy from the previous measurement. Filter this table to show all of the entries that have experienced a drop in life expectancy.
Hint: you might find the lag() or diff() functions useful.

```{r new variable}
gapminder %>%
  mutate(lifeExp_change=diff(lifeExp)

```

